---
title: "Project Sample ~ CL"
author: "JWK"
date: "7/2/2021"
output: html_document
---

```{r include=FALSE}
setwd("C:/Users/jonat/Desktop/machine-learning-toolbox/07_project_samples/project.sample.classification")
```

# - Load the libraries 

These are required packages for the project.sample. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(remotes)
library(corrplot)
library(cowplot)
library(outlieR)
library(magrittr)
library(NoiseFiltersR)
library(caret)
library(ggplot2)
library(AppliedPredictiveModeling)
```

# - Import the Original Data 

```{r echo=TRUE, message=FALSE, warning=FALSE}
source("scripts/import_data.r")
```

## - Exploration of the Original Data 

The following section will detail the original data, including a statistical summary, scatterplot grid, histogram grid, barplot grid, line plot grid, and a correlation plot. 
The original data's variables are set as followed:

```{r echo=TRUE, message=FALSE, warning=FALSE}
x1 = data$Sepal.Length
x2 = data$Sepal.Width
x3 = data$Petal.Length
x4 = data$Petal.Width
x5 = data$Species
d1 = data[,1:4]
```

### - Statistical Summary of the Original Data 

```{r echo=TRUE, message=FALSE, warning=FALSE}
summary(data)
```

### - Scatter Plot Grid of the Original Data 

```{r echo = F, fig.cap = 'ScatterPlot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/mainscatter_O.png")
grid.raster(img)
```

### - Histogram Plot Grid of the Original Data 

```{r echo = F, fig.cap = 'Histogram Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/hist_O.png")
grid.raster(img)
```

### - Bar Plot Grid of the Original Data 

```{r echo = F, fig.cap = 'Bar Plot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/bar_O.png")
grid.raster(img)
```

### - Line Plot Grid of the Original Data 

```{r echo = F, fig.cap = 'Line Plot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/line_O.png")
grid.raster(img)
```

### - Correlation Plot of the Original Data 

```{r echo = F, fig.cap = 'Bar Plot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/cor_O.png")
grid.raster(img)
```

# - Import the Pre-processed Data

```{r echo=TRUE, message=FALSE, warning=FALSE}
source("scripts/import_prep.r")
```

## - Exploration of the Pre-processed Data 

The following section will detail the pre-processed data, including a statistical summary, scatterplot grid, histogram grid, barplot grid, line plot grid, and a correlation plot. 
The pre-processed data's variables are set as followed:

```{r echo=TRUE, message=FALSE, warning=FALSE}
x1 = prep$Sepal.Length
x2 = prep$Sepal.Width
x3 = prep$Petal.Length
x4 = prep$Petal.Width
x5 = prep$x5
d1 = prep[,1:4]
```

### - Statistical Summary of the Pre-processed Data

```{r echo=TRUE, message=FALSE, warning=FALSE}
summary(prep)
```

### - Scatter Plot Grid of the Pre-processed Data 

```{r echo = F, fig.cap = 'ScatterPlot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/mainscatter_P.png")
grid.raster(img)
```

### - Histogram Plot Grid of the Pre-processed Data 

```{r echo = F, fig.cap = 'Histogram Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/hist_P.png")
grid.raster(img)
```

### - Bar Plot Grid of the Pre-processed Data 

```{r echo = F, fig.cap = 'Bar Plot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/bar_P.png")
grid.raster(img)
```

### - Line Plot Grid of the Pre-processed Data 

```{r echo = F, fig.cap = 'Line Plot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/line_P.png")
grid.raster(img)
```

### - Correlation Plot of the Pre-processed Data 

```{r echo = F, fig.cap = 'Bar Plot Grid'}
library(png)
library(grid)
img <- readPNG("./pictures/cor_P.png")
grid.raster(img)
```

# - Modeling using a Decision Tree

The trained model was able to produce the following results:

## - Trained Model Results 

```{r echo=FALSE, message=FALSE, warning=FALSE}
model <- readRDS("./states/dt_model.rds")
print(model)
```

## - Trained Model Plot 

```{r echo = F, fig.cap = 'Trained Model Plot'}
library(png)
library(grid)
img <- readPNG("./pictures/dt_model_plot.png")
grid.raster(img)
```

## - Trained Models Evaluation 

```{r echo=FALSE, message=FALSE, warning=FALSE}
index = createDataPartition(prep[,1], p =0.80, list = FALSE)
valid = prep[-index,]
data.pred = predict(model, newdata = valid)
cm = confusionMatrix(as.factor(data.pred), reference = as.factor(valid$x5), mode = "prec_recall")
print(cm)
```

