---
title: Machine Learning Toolbox Regression Report
author: "J.W.Korn"
date: "`r Sys.Date()`"
output: html_document
---

## Load Libraries, Import Data, and Correlation {.tabset}

### Load Libraries 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(caret)
library(rattle)
library(party)
library(ggpubr)
library(magrittr)
library(datasets)
library(mice)
library(outlieR)
library(AppliedPredictiveModeling)
```

Here is the list of all the required libraries. 
Download all the libraries before running the code. 

### Import Data 
```{r echo=FALSE, message=FALSE, warning=FALSE}
data = airquality
str(data)
```

### Remove Unnecessary Variables 
```{r}
dataset = data[,-c(5,6)]
str(dataset)
```

### Missing and Outliers  {.tabset}
```{r echo=FALSE, message=FALSE, warning=FALSE}
apply(dataset,2,function(x) sum(is.na(x)))#Check for Missing...
dataset = na.omit(dataset)
summary(dataset)
```

### Impute for Missing or NA Values
```{r echo=FALSE, message=FALSE, warning=FALSE}
missing = dataset %>% mice::mice(m=5,maxit=50,meth="sample",seed=500,print = FALSE)
missing <- mice::complete(missing, action=as.numeric(2))
dataset = na.omit(missing)
print(str(dataset))
```

### Impute Outliers 
```{r echo=FALSE, message=FALSE, warning=FALSE}
data = dataset %>% outlieR::impute(flag = NULL, fill = "mean", 
                                level = 0.1, nmax = NULL,
                                side = NULL, crit = "lof", 
                                k = 5, metric = "euclidean", q = 3)
print(str(dataset))
```

### Normalize the Data 
```{r echo=FALSE, message=FALSE, warning=FALSE}
preProClean <- preProcess(x = dataset, method = c("scale", "center"))
dataset <- predict(preProClean, dataset %>% na.omit)
print(str(dataset))
```


### Scatterplot Matrix
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(GGally)
sp = ggpairs(dataset)
sp
```

### Scatter Plot
```{r echo=FALSE, message=FALSE, warning=FALSE}
plot(dataset[, "Solar.R"], main = "Solar R")
```


### Split the Data 
```{r}
index = createDataPartition(dataset[,1], p =0.80, list = FALSE)
training = dataset[index,]
valid = dataset[-index,]
```


## Train Models {.tabset}

### Modeling Parameters 
```{r echo=TRUE, message=FALSE, warning=FALSE}
control <- trainControl(method="cv", number=10)
metric <- c("RMSE")
set.seed(7)
```

### Train Models

- Decision Tree 
- Random Forest 

```{r echo=TRUE, message=FALSE, warning=FALSE}
set.seed(7)
fit.dt <- train(Solar.R~., data = training, method="rpart", metric=metric, trControl=control) #decision tree
set.seed(7)
fit.rf <- train(Solar.R~., data = training, method="rf", metric=metric, trControl=control, importance=T) #random forest
```

### Summary Results 
```{r echo=FALSE, message=FALSE, warning=FALSE}
results = resamples(list(dt= fit.dt, rf = fit.rf))
summary(results)
```

### Summary Plots
```{r echo=FALSE, message=FALSE, warning=FALSE}
dp = dotplot(results)
dp
bwp = bwplot(results)
bwp
```

## Decision Tree Results {.tabset}

### Summary Results
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(fit.dt$resample)
```

### Mean of Different Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.dt$results)
```

### Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.dt$resample)
```

### Predictions Results 
```{r echo=FALSE, message=FALSE, warning=FALSE}
predictions = predict(fit.dt, valid)
postResample(pred = predictions, obs = valid$Solar.R)
```

### Variable Importance
```{r echo=FALSE, message=FALSE, warning=FALSE}
importance <- varImp(fit.dt, scale=TRUE)
# summarize importance
print(importance)
# plot importance
vi.dt = plot(importance)
vi.dt
```


## Random Forest Results {.tabset}

### Summary Results
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(fit.rf$resample)
```

### Mean of Different Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.rf$results)
```

### Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.rf$resample)
```

### Predictions
```{r echo=FALSE, message=FALSE, warning=FALSE}
predictions = predict(fit.rf, valid)
postResample(pred = predictions, obs = valid$Solar.R)
```

### Variable Importance
```{r echo=FALSE, message=FALSE, warning=FALSE}
importance <- varImp(fit.rf, scale=TRUE)
# summarize importance
print(importance)
# plot importance
vi.rf = plot(importance)
vi.rf
```


## Project Summary Plots {.tabset}

### Scatterplot Matrix
```{r echo=FALSE, message=FALSE, warning=FALSE}
sp
```

### Training Summary Plots
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(dp, bwp, 
          labels = c("Training Models Summary Results"),
          ncol = 2, nrow = 1)
```

### VI Comparison Plot
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(vi.dt, vi.rf, 
          labels = c("DT Original", 
                     "RF Original"),
          ncol = 2, nrow = 2)
```
