---
title: Machine Learning Sample
  Learning
author: "HU - J.Korn"
date: "`r Sys.Date()`"
output: html_document
---
```{r message=FALSE, warning=FALSE, include=FALSE}
setwd("C:/Users/Jonathan Korn/Desktop/cu_hsp-learning/09_reporting")
```

## Load Libraries, Import Data, and Correlation {.tabset}

### Load Libraries 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(caret)
library(rattle)
library(party)
library(ggpubr)
library(quantmod)
```

Here is the list of all the required libraries. 
Download all the libraries before running the code. 

### Import Data 
```{r echo=FALSE, message=FALSE, warning=FALSE}
start <- as.Date(Sys.Date()-(365*5)) # start date. 
end <- as.Date(Sys.Date()) # current date. 
getSymbols("GOOG", src = "yahoo", from = start, to = end) # Feed symbol from s&p500. 
data = data.frame(GOOG) # call xts object into env. 
colnames(data) = c("Open", "High", "Low", "Close", "Volume", "Adjusted") # set column names in xts object. 
str(data)
```

### Remove Unnecessary Variables 
```{r}
dataset = data[,-c(5,6)]
str(dataset)
```

### Check for Missing or NA Values {.tabset}
```{r echo=FALSE, message=FALSE, warning=FALSE}
apply(dataset,2,function(x) sum(is.na(x)))#Check for Missing...
dataset = na.omit(dataset)
summary(dataset)
```

### Scatterplot Matrix
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(GGally)
sp = ggpairs(dataset)
sp
```

## Train Models {.tabset}

### Modeling Parameters 
```{r echo=TRUE, message=FALSE, warning=FALSE}
control <- trainControl(method="cv", number=10)
metric <- c("RMSE")
set.seed(7)
```

### Train Models

- Regression 
- Random Forest 

```{r echo=TRUE, message=FALSE, warning=FALSE}
set.seed(7)
fit.OLS <- train(Close~., data = dataset, method="lm", metric=metric, trControl=control) #OLS
set.seed(7)
fit.rf <- train(Close~., data = dataset, method="rf", metric=metric, trControl=control, #random forest 
               importance=T)
```

### Summary Results 
```{r echo=FALSE, message=FALSE, warning=FALSE}
results = resamples(list(OLS= fit.OLS, rf = fit.rf))
summary(results)
```

### Summary Plots
```{r echo=FALSE, message=FALSE, warning=FALSE}
dp = dotplot(results)
dp
bwp = bwplot(results)
bwp
```

## Regression Results {.tabset}

### Summary Results
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(fit.OLS$resample)
```

### Mean of Different Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.OLS$results)
```

### Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.OLS$resample)
```

### Predictions Results 
```{r echo=FALSE, message=FALSE, warning=FALSE}
predictions = predict(fit.OLS, dataset[,-4])
postResample(pred = predictions, obs = dataset$Close)
```

### Variable Importance
```{r echo=FALSE, message=FALSE, warning=FALSE}
importance <- varImp(fit.OLS, scale=TRUE)
# summarize importance
print(importance)
# plot importance
vi.OLS = plot(importance)
vi.OLS
```


## Random Forest Results {.tabset}

### Summary Results
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(fit.rf$resample)
```

### Mean of Different Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.rf$results)
```

### Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.rf$resample)
```

### Predictions
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Estimate Skill of a model using the Validation Set
predictions = predict(fit.rf, dataset[,-4])
postResample(pred = predictions, obs = dataset$Close)
```

### Variable Importance
```{r echo=FALSE, message=FALSE, warning=FALSE}
importance <- varImp(fit.rf, scale=TRUE)
# summarize importance
print(importance)
# plot importance
vi.rf = plot(importance)
vi.rf
```


## Regression with Significant Variables - Results {.tabset}

### Train on Significant Variables
```{r echo=TRUE, message=FALSE, warning=FALSE}
fit.OLS.r <- train(Close~High+Low, data = dataset, method="lm", metric=metric, trControl=control, #random forest 
               importance=T)
```

### Summary Results
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(fit.OLS.r$resample)
```

### Mean of Different Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.OLS.r$results)
```

### Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.OLS.r$resample)
```

### Predictions
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Estimate Skill of a model using the Validation Set
predictions = predict(fit.OLS.r, dataset[,-4])
postResample(pred = predictions, obs = dataset$Close)
```

### Variable Importance
```{r echo=FALSE, message=FALSE, warning=FALSE}
importance <- varImp(fit.OLS.r, scale=TRUE)
# summarize importance
print(importance)
# plot importance
vi.OLS.r = plot(importance)
vi.OLS.r
```


### Training on the Whole Dataset
```{r echo=FALSE, message=FALSE, warning=FALSE}
lm = coef(lm(Close~High + Low, data = dataset)) #Same as running using caret with no control parameters...
summary(lm)
```

### Plot
```{r echo=FALSE, message=FALSE, warning=FALSE}
fmp.1 = plot(fit.OLS.r$finalModel,1) #same as the above,,,
fmp.2 = plot(fit.OLS.r$finalModel,2)
fmp.3 = plot(fit.OLS.r$finalModel,3)
fmp.4 = plot(fit.OLS.r$finalModel,4)
```

## Random Forest With Significant Variables - Results {.tabset}

### Train on Significant Variables
```{r echo=TRUE, message=FALSE, warning=FALSE}
fit.rf.r <- train(Close~High + Low, data = dataset, method="rf", metric=metric, trControl=control, #random forest 
               importance=T)
```

### Summary Results
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(fit.rf.r$resample)
```

### Mean of Different Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.rf.r$results)
```

### Resamples
```{r echo=FALSE, message=FALSE, warning=FALSE}
print(fit.rf.r$resample)
```

### Predictions 
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Estimate Skill of a model using the Validation Set
predictions = predict(fit.rf.r, dataset[,-4])
postResample(pred = predictions, obs = dataset$Close)
```

### Variable Importance
```{r echo=FALSE, message=FALSE, warning=FALSE}
importance <- varImp(fit.rf.r, scale=TRUE)
# summarize importance
print(importance)
# plot importance
vi.rf.r = plot(importance)
vi.rf.r
```

### Forest Plot
```{r echo=FALSE, message=FALSE, warning=FALSE}
fmp2 = plot(fit.rf.r$finalModel)
```

## Project Summary Plots {.tabset}

### Scatterplot Matrix
```{r echo=FALSE, message=FALSE, warning=FALSE}
sp
```

### Training Summary Plots
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(dp, bwp, 
          labels = c("Training Models Summary Results"),
          ncol = 2, nrow = 1)
```

### VI Comparison Plot
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(vi.OLS, vi.rf, vi.OLS.r, vi.rf.r, 
          labels = c("LM Original", 
                     "RF Original", 
                     "LM Significant", 
                     "RF Significant"),
          ncol = 2, nrow = 2)
```
